### 2023-06-06 snp_effects/cauchy_distrn.Rmd
snp effects code

Let's check this works for the Cauchy.

```{r fake_data}
n <- 1e3
gwas <- data.frame(minor_AF=runif(n),
                   beta=rcauchy(n)/100)

min_num <- 20
max_prop <- 0.5
```

Now let's let $(P, Z)$ be a random draw from this table, all $Z > 0$,
and $X = Z$ with probability $P$, and $X=0$ otherwise.
Here's the complementary cumulative distribution function of $X$,
i.e., if $N$ is the number of rows in the table,
$$\begin{aligned}
   \mathbb{P}\{X > x \}  
   &= \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i > x} \\
   &= \mathbb{P}\{X > 0\} - \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i \le x} \\
   &= \frac{1}{N}\sum_i P_i - \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i \le x} .
\end{aligned}$$

```{r plotit}
estim_alpha <- function (gwas, num_breaks=1001) {
    xx <- c(seq(0, 2 * quantile(abs(gwas$beta), 0.99), length.out=num_breaks), Inf)
    disc_beta <- cut(abs(gwas$beta), xx)
    num_points <- table(disc_beta)
    dpx <- tapply(gwas$minor_AF, disc_beta, sum)
    dpx[is.na(dpx)] <- 0
    p_zero <- mean(gwas$minor_AF)
    tail_prob <- p_zero - cumsum(dpx) / nrow(gwas)
    tail_nums <- nrow(gwas) - cumsum(num_points)
    x <- xx[-1]
    xlm <- lm(log(tail_prob) ~ log(x),
              subset=(is.finite(log(tail_prob))
                      & tail_nums > min_num
                      & tail_nums < max_prop * nrow(gwas)))
    return(xlm)
}

xlm <- estim_alpha(gwas)
logx <- xlm$model[[1]]
logy <- xlm$model[[2]]
plot(logx, logy, ylab='log(prob > x)', xlab='log(x)')
abline(coef(xlm))

summary(xlm)
```

## Power

Huh, how good of an estimator is this?
```{r boot}
boots <- replicate(200, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rcauchy(n)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(boots[2,], breaks=30)
abline(v=-1, col='red')
```
Looks ok.

## Gaussian

And, what if we give this Gaussian data?
```{r gboot}
boots <- replicate(200, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rnorm(n)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(boots[2,], breaks=30)
abline(v=-1, col='red')
```
Looks ok.